Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2022-06-09 E - Tutorial Log-Normal Model Rstan/Online Supplement E - Bayesian Hierarchical Modeling in rstan - Log Model.Rmd
All documents except .rds

---
title             : "Online Supplement C: Bayesian Hierarchical Modeling in rstan"
author            : "Myrthe Veenman" 
bibliography      : ["r-references.bib"]
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}

@@

|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan]. In addition, it contains information about the Number Classification Task that could not be included in the paper. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. This tutorial will cover the following:

@@

|    To fit a BHM in *rstan*, we use the *stan* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- stan(file = "./myModel.stan",  # Stan file with model  
                  data = myData_list,       # List with observed data and constants
                  iter = 4000,              # Number (Nr.) of iterations per chain
                  chains = 4,               # Nr. of chains 
                  warmup = 1000)            # Nr. iterations for warmup per chain
```
\normalsize

@@

|  The stan file `myModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model. It contains descriptions for every line of code. The stan file is divided in three sections (in our case):

@@

\scriptsize 
```{r priors}
### Set the priors and add them to the list 
a <- 0.5        # mean for mu, mu is the mean of theta  
b <- 1          # variance for mu 
c <- 3          # mean for sigma 
d <- 0.7        # variance for sigma, sigma is the variance of observations  
e <- 3          # mean for g, g is the variance of theta   
f <- 0.7        # variance for g

#### Priors for beta 
a2 <- 0        # mean for mean of beta
b2 <- 0.09     # variance for mean of beta, used to be 1 
e2 <- 3        # mean for variance of beta, used to be .7, adjusted 
f2 <- 0.5      # variance for variance of beta, used to be .2., adjusted

#### Prior for deltas, same for every delta 
a3 <- 0        # mean for mean of delta 
b3 <- 0.09     # sd for mean of delta, 0.5, 0.3  
dd3 <- 3       # mean for variance of delta, adjusted
f3 <- 0.5      # variance for variance of delta, adjusted

@@

datalistpp3 <- list()
for (i in 1:1000){
  # Hyperpriors: These are not individual specific right? So only sample 1? 
  mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
  vargammapp <- rinvgamma(1, e, f)
  mubetapp <- rnorm(1, a2, sqrt(b2))
  varbetapp <- rinvgamma(1, e2, f2)
  mudeltapp <- rnorm(1, a3, sqrt(b3))
  vardeltapp <- rinvgamma(1, dd3, f3)

@@

  # Or start for loop here? Or do we want to vary persons as well? 
  # data
  Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                 ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
                 wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
               sqrt(sigmapp))
  datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
  datalistpp3[[i]] <- datartpp
}

# Add datasets under each other 
bigdatapp3_rstan <- do.call(rbind, datalistpp3)

@@

# Create figure 
priorpredictionplotnew_2 <- priorpredfigurecreate(bigdatapp3_rstan)

@@

\scriptsize 
``` {r modelfit, eval = FALSE}
library(rstan) 
model_fit <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
                  data = myData_list, 
                  iter = 4000, 
                  chains = 4,
                  warmup = 1000, 
                  cores = 4)
```
\normalsize

We have already fitted the model and saved it in the R object `model_fit`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimates.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
model_fit <- readRDS(file = "myPath/rstan_model_fit.rds")
```
\normalsize 
```{r loadmodelfit2, echo = FALSE}
setwd('..')
model_fit <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
```

@@

|    *Brms* offers the *launch_shinystan* function which will load a shiny app with the results of the model fit. It is very extensive. It, for instance, includes model diagnostics as trace plots, posterior plots, prediction plots and way more. 
&nbsp;

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(model_fit)
```
\normalsize

@@

|    The *summary* function provides an overview of the model fit. From this function, we can get a lot of information. As the output is structured in a specific way, it sometimes is hard to figure out how to isolate specific outcomes. [This document](https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html) nicely explains how to obtain specific output from the model fit.

@@

|    The trace plot shows whether the posterior distribution of the parameter has converged. With the package's *plot* function it is possible to obtain the trace plots per parameter by setting the option *plotfun* to *trace*. By default, it will show the trace plots of the first ten parameters. With the *pars* function you can specify for which parameters you want to inspect the trace plots. The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, if you would like to see that as well you have to set the option *inc_warmup* to *TRUE*. As the *plot* option returns a *ggplot* object, it is possible to manually adjust the plot using *ggplot* functions. We created the function *traceplotfunc* to obtain the trace plot for the parameter you are interested in. In the figure below, we show the trace plots for all the general effects. 
&nbsp;

@@

# Trace plot for parameter mu4
trdif2rstan <- plot(model_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c("mu4"))     # Select parameter, 
#it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

# In case you want to see the warmup period as well 
trdif2rstanwarmup <- plot(model_fit,    # Model fit 
                          plotfun = "trace",  # Specify that you want trace plots 
                          pars = c("mu4"),    # Select parameter
                          inc_warmup = TRUE)  # Show warmup period as well

@@

# Trace plot for parameter mu4
trdif2rstan <- plot(model_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c(parameter))     # Select parameter, 
# it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

@@

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
plot(model_fit, plotfun= "rhat")

# Plot frequency of the ratio 
plot(model_fit, plotfun = "ess")
```
\normalsize

@@

\scriptsize 
```{r rhatneff, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8, eval = FALSE, echo = FALSE}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(model_fit)$summary[,"Rhat"]

# Obtain number of effective samples for every paramter from model fit 
neffrstan <- summary(model_fit)$summary[,"n_eff"]

@@

\scriptsize 
```{r rhatneff2, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(model_fit)$summary[,"Rhat"]
fixedrhatrstan <- rhatrstan[1:13]  # for general effects 
deltameanrrstan <- mean(fixedrhatrstan[3:6])  # calculate the mean for delta parameters

# Obtain number of effective samples for every parameter from model fit 
neffrstan <- summary(model_fit)$summary[,"n_eff"]
fixedneffrstan <- neffrstan[1:13]  # for general effects 
fixedneffrstanmean <- mean(fixedneffrstan[3:6])  # calculate the mean for delta parameters

@@

|    With the *plot* function, the posterior distribution per parameter can be displayed. 
&nbsp;

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
plot(model_fit, 
     plotfun = "dens")  
# without specifying pars, the posterior distributions of the first ten parameters are displayed 
```
\normalsize

|    Another possibility is to use the *bayesplot* package to plot all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
&nbsp;

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
postc <- as.array(model_fit)  # change structure of model fit

@@

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(model_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

@@

|    First, we can show a table with the posterior distribution estimates of the variance parameters. This indicates the variation of the general effect. We called our variances *g* (*g* until *g6*). 
&nbsp;

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
fitrstanhiertable2 <- summary(model_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

@@

# Extract the parameter estimates using the names from before 
gamstan <- as.data.frame(summary(model_fit, 
                                 pars = c( gamnamesstan))$summary)

# Extract the general effect estimate of the parameter 
gamstanmu <- as.data.frame(summary(model_fit, 
                                   pars = c("mu"))$summary)

@@

# Beta  
betnamesstan <- paste("betai", "[",1:52,"]", sep="")
betstan <- as.data.frame(summary(model_fit, pars = c( betnamesstan))$summary)
betstanmu <- as.data.frame(summary(model_fit, pars = c("mu2"))$summary)
betstan$parameters <- c(betnamesstan)

@@

# Delta1 
del1namesstan <- paste("delta1", "[",1:52,"]", sep="")
delt1stan <- as.data.frame(summary(model_fit, pars = c( del1namesstan))$summary)
del1stanmu <- as.data.frame(summary(model_fit, pars = c("mu3"))$summary)
delt1stan$parameters <- c(del1namesstan)

@@

# Delta 2 
del2namesstan <- paste("delta2", "[",1:52,"]", sep="")
delt2stan <- as.data.frame(summary(model_fit, pars = c( del2namesstan))$summary)
del2stanmu <- as.data.frame(summary(model_fit, pars = c("mu4"))$summary)
delt2stan$parameters <- c(del2namesstan)

@@

# Delta 3 
del3namesstan <- paste("delta3", "[",1:52,"]", sep="")
delt3stan <- as.data.frame(summary(model_fit, pars = c( del3namesstan))$summary)
del3stanmu <- as.data.frame(summary(model_fit, pars = c("mu5"))$summary)
delt3stan$parameters <- c(del3namesstan)

@@

# Delta 4 
del4namesstan <- paste("delta4", "[",1:52,"]", sep="")
delt4stan <- as.data.frame(summary(model_fit, pars = c( del4namesstan))$summary)
del4stanmu <- as.data.frame(summary(model_fit, pars = c("mu6"))$summary)
delt4stan$parameters <- c(del4namesstan)

@@

### Obtain all individual estimates delta 
hier_modeld <- model_fit  # model fit

@@

|    In case we have different models, we can compare them to check under which model the data is most likely. This can be done with the Bayes factor. In *rstan* the Bayes factor can be obtained by using the *bridgesampling* package. To be able to use this package, it is important that the *.stan* file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).   
|    First, we fit all the models (except for the full model that we already fitted. 
&nbsp;

\scriptsize 
```{r modelcomparison, eval = FALSE, echo = TRUE}
# Fit the other models 
## Null model 
BFnullmodel <- stan(file = "myPath/rewritestanmodel3h0.stan", 
                    data = myData_list, 
                    iter = 4000, 
                    chains = 4,  
                    warmup = 1000, 
                    cores = 4)

## Side model 
BFsidemodel <- stan(file = "myPath/rewritestanmodel3hside.stan", 
                       data = myData_list, 
                       iter = 4000, 
                       chains = 4, 
                       warmup = 1000, 
                       cores = 4)

## Digit model 
BFdigitmodel <- stan(file = "myPath/rewritestanmodel3hdigit.stan", 
                         data = myData_list, 
                         iter = 4000, 
                         chains = 4, 
                         warmup = 1000, 
                         cores = 4)
```
\normalsize

@@

\scriptsize 
```{r bfpart2, eval = FALSE, echo = TRUE}
# Bridge sampling 
# library(bridgesampling)
bridge_Hnull <- bridge_sampler(BFnullmodel)
bridge_Hside <- bridge_sampler(BFsidemodel)
bridge_Hdigit <- bridge_sampler(BFdigitmodel)
bridge_Hfull <- bridge_sampler(model_fit)  # in case you just fitted the model yourself

@@

### Fit model again but do not run any chains 
new_mod <- stan(file = "myPath/rewritestanmodel3_julia.stan", 
                data = myData_list, 
                chains = 0)

### Now do the bridgesampling 
bridge_Hfull <- bridge_sampler(model_fit, new_mod)
```
\normalsize

@@

# Obtain BF (in this way evidence in favor of the null hypothesis)
bridgesampling::bf(bridge_Hnull, bridge_Hfull)   # Null against full

# Obtain BF (in this way evidence in favor of the full hypothesis)
bridgesampling::bf(bridge_Hfull, bridge_Hnull)   # Full against null
bridgesampling::bf(bridge_Hfull, bridge_Hside)   # Full against side 
bridgesampling::bf(bridge_Hfull, bridge_Hdigit)  # Full against digit 
```
\normalsize

>>>>

---
title             : "Online Supplement E: Bayesian Hierarchical Modeling in rstan - Log Model"
author            : "Myrthe Veenman" 
bibliography      : ["r-references.bib"]
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \usepackage{graphicx}
   - \usepackage{float}

@@

|    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `rstan` [@R-rstan]. In addition, it contains information about the Number Classification Task that could not be included in the paper. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. This tutorial will cover the following:

@@

|    To fit a log-normal BHM in *rstan*, we use the *stan* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- stan(file = "./myLogModel.stan",  # Stan file with model  
                  data = myData_list,       # List with observed data and constants
                  iter = 4000,              # Number (Nr.) of iterations per chain
                  chains = 4,               # Nr. of chains 
                  warmup = 1000)            # Nr. iterations for warmup per chain
```
\normalsize

@@

|  The stan file `myLogModel.stan` is used to fit the full model and can be found under R objects > rstan > linear model. It contains descriptions for every line of code. The stan file is divided in three sections (in our case):

@@

\scriptsize 
```{r priors}
### Set the priors and add them to the list 
a <- -0.5        # mean for mu, mu is the mean of theta  
b <- 1          # variance for mu 
c <- 3          # mean for sigma 
d <- 0.3        # variance for sigma, sigma is the variance of observations  
e <- 3          # mean for g, g is the variance of theta   
f <- 0.3        # variance for g

#### Priors for beta 
a2 <- 0        # mean for mean of beta
b2 <- 0.005     # variance for mean of beta, used to be 1 
e2 <- 3        # mean for variance of beta, used to be .7, adjusted 
f2 <- 0.01      # variance for variance of beta, used to be .2., adjusted

#### Prior for deltas, same for every delta 
a3 <- 0        # mean for mean of delta 
b3 <- 0.005     # sd for mean of delta, 0.5, 0.3  
dd3 <- 3       # mean for variance of delta, adjusted
f3 <- 0.01      # variance for variance of delta, adjusted

@@

### Set the priors and add them to the list 
a <- -0.5  # mean for mu, mu is the mean of theta  
b <- 1  # variance for mu 
c <- 3  # mean for sigma 
d <- .3  # variance for sigma, sigma is the variance of observations  
e <- 3  # mean for g, g is the variance of theta   
f <- .3  # variance for g

#### Priors for beta 
a2 <- 0  # mean for mean of beta
b2 <- 0.005  # sd for mean of beta, used to be 1 
e2 <- 3  # mean for sd of beta, used to be .7 
f2 <- 0.01  # sd for sd of beta, used to be .2.

#### Prior for deltas, same for every delta 
a3 <- 0  # mean for mean of delta 
b3 <- 0.005  # sd for mean of delta, 0.005 pretty good, 0,01 worse 
dd3 <- 3  # mean for sd of delta 
f3 <- 0.01  # sd for sd of delta, 0.01 good, 0.05 worse

datalistpp3 <- list()
delta1pp2 <- list()
gammaipp2 <- list()
betaipp2 <- list()
sigmapp2 <- c()
for (i in 1:1000){  # with 10000 my computer cannot handle it 
  # Hyperpriors: These are not individual specific right? So only sample 1? 
  mugammapp <- rnorm(1, -0.5, sqrt(1))
  vargammapp <- rinvgamma(1, e, f)
  mubetapp <- rnorm(1, a2, sqrt(b2))
  varbetapp <- rinvgamma(1, e2, f2)
  mudeltapp <- rnorm(1, a3, sqrt(b3))
  vardeltapp <- rinvgamma(1, dd3, f3)

@@

  # Or start for loop here? Or do we want to vary persons as well? 
  # data
  Ypp <- rlnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                  ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
                  wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
                sqrt(sigmapp))
  gammaipp2[[i]] <- gammaipp
  betaipp2[[i]] <- betaipp
  delta1pp2[[i]] <- delta1pp
  sigmapp2[i] <- sigmapp
  datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
  datalistpp3[[i]] <- datartpp
}

# Add datasets under each other 
bigdatapp3 <- do.call(rbind, datalistpp3)

@@

# Create figure 
priorpredictionplotnew_2 <- priorpredfigurecreate(bigdatapp3)

@@

\scriptsize 
``` {r modelfit, eval = FALSE}
library(rstan) 
model_fit <- stan(file = "myPath/myLogModel.stan", 
                  data = myData_list, 
                  iter = 4000, 
                  chains = 4,
                  warmup = 1000, 
                  cores = 4)
```
\normalsize

We have already fitted the model and saved it in the R object `logmodel_fit`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimates.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
logmodel_fit <- readRDS(file = "myPath/rstan_logmodel_fit.rds")
```
\normalsize 
```{r loadmodelfit2, echo = FALSE}
setwd('..')
logmodel_fit <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
```

@@

|    `Brms` offers the `launch_shinystan` function which will load a shiny app with the results of the model fit. It is very extensive. It, for instance, includes model diagnostics as trace plots, posterior plots, prediction plots and way more. 
&nbsp;

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(logmodel_fit)
```
\normalsize

@@

|    The `summary` function provides an overview of the model fit. From this function, we can get a lot of information. As the output is structured in a specific way, it sometimes is hard to figure out how to isolate specific outcomes. [This document](https://cran.r-project.org/web/packages/rstan/vignettes/stanfit-objects.html) nicely explains how to obtain specific output from the model fit.

@@

|    The trace plot shows whether the posterior distribution of the parameter has converged. With the package's `plot` function it is possible to obtain the trace plots per parameter by setting the option `plotfun` to `trace`. By default, it will show the trace plots of the first ten parameters. With the `pars` function you can specify for which parameters you want to inspect the trace plots. The trace plot will automatically remove the warm-up (burn-in) period from the chain (which is desirable), however, if you would like to see that as well you have to set the option `inc_warmup` to `TRUE`. As the `plot` option returns a `ggplot` object, it is possible to manually adjust the plot using `ggplot` functions. We created the function `traceplotfunc` to obtain the trace plot for the parameter you are interested in. In the figure below, we show the trace plots for all the general effects. 
&nbsp;

@@

# Trace plot for parameter mu4
trdif2rstan <- plot(logmodel_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c("mu4"))     # Select parameter, 
#it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

# In case you want to see the warmup period as well 
trdif2rstanwarmup <- plot(logmodel_fit,    # Model fit 
                          plotfun = "trace",  # Specify that you want trace plots 
                          pars = c("mu4"),    # Select parameter
                          inc_warmup = TRUE)  # Show warmup period as well

@@

# Trace plot for parameter mu4
trdif2rstan <- plot(logmodel_fit,     # Model fit 
                    plotfun = "trace",   # Specify that you want trace plots 
                    pars = c(parameter))     # Select parameter, 
# it is also possible to select more than one at the same time, such as c("mu2", "mu3","mu4")

@@

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
plot(logmodel_fit, plotfun= "rhat")

# Plot frequency of the ratio 
plot(logmodel_fit, plotfun = "ess")
```
\normalsize

@@

\scriptsize 
```{r rhatneff, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8, eval = FALSE, echo = FALSE}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(logmodel_fit)$summary[,"Rhat"]

# Obtain number of effective samples for every paramter from model fit 
neffrstan <- summary(logmodel_fit)$summary[,"n_eff"]

@@

\scriptsize 
```{r rhatneff2, fig.cap="The frequency of the $\\hat{R}$ statistic and the number of effective samples. A: The frequency of the $\\hat{R}$. B: The frequency of the number of effective samples where the dashed lines represent the total number of iterations.", fig.align = "left", fig.height = 4, fig.width = 8}
# Obtain rhat for every parameter from model fit 
rhatrstan <- summary(logmodel_fit)$summary[,"Rhat"]
fixedrhatrstan <- rhatrstan[1:13]  # for general effects 
deltameanrrstan <- mean(fixedrhatrstan[3:6])  # calculate the mean for delta parameters

# Obtain number of effective samples for every parameter from model fit 
neffrstan <- summary(logmodel_fit)$summary[,"n_eff"]
fixedneffrstan <- neffrstan[1:13]  # for general effects 
fixedneffrstanmean <- mean(fixedneffrstan[3:6])  # calculate the mean for delta parameters

@@

|    With the `plot` function, the posterior distribution per parameter can be displayed. 
&nbsp;

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
plot(logmodel_fit, 
     plotfun = "dens")  
# without specifying pars, the posterior distributions of the first ten parameters are displayed 
```
\normalsize

|    Another possibility is to use the `bayesplot` package to plot all the posterior densities together in one plot. This makes it easier to compare the posterior distributions of each parameter. 
&nbsp;

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
postc <- as.array(logmodel_fit)  # change structure of model fit

@@

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitrstanhiertable <- summary(logmodel_fit,  # model fit 
                             # Select parameters 
                             pars = c("mu", "mu2", "mu3", "mu4", "mu5", 
                                      "mu6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

@@

|    First, we can show a table with the posterior distribution estimates of the variance parameters. This indicates the variation of the general effect. We called our variances `g` (`g` until `g6`). 
&nbsp;

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
fitrstanhiertable2 <- summary(logmodel_fit,   # model fit 
                              # Select variance parameters (we called them g)
                              pars = c("g", "g2", "g3", "g4", "g5", 
                                       "g6"))$summary[,c("mean", "se_mean", 
                                                         "2.5%", "97.5%", 
                                                         "n_eff", "Rhat")]

@@

# Extract the parameter estimates using the names from before 
gamstan <- as.data.frame(summary(logmodel_fit, 
                                 pars = c( gamnamesstan))$summary)

# Extract the general effect estimate of the parameter 
gamstanmu <- as.data.frame(summary(logmodel_fit, 
                                   pars = c("mu"))$summary)

@@

# Beta  
betnamesstan <- paste("betai", "[",1:52,"]", sep="")
betstan <- as.data.frame(summary(logmodel_fit, pars = c( betnamesstan))$summary)
betstanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu2"))$summary)
betstan$parameters <- c(betnamesstan)

@@

# Delta1 
del1namesstan <- paste("delta1", "[",1:52,"]", sep="")
delt1stan <- as.data.frame(summary(logmodel_fit, pars = c( del1namesstan))$summary)
del1stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu3"))$summary)
delt1stan$parameters <- c(del1namesstan)

@@

# Delta 2 
del2namesstan <- paste("delta2", "[",1:52,"]", sep="")
delt2stan <- as.data.frame(summary(logmodel_fit, pars = c( del2namesstan))$summary)
del2stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu4"))$summary)
delt2stan$parameters <- c(del2namesstan)

@@

# Delta 3 
del3namesstan <- paste("delta3", "[",1:52,"]", sep="")
delt3stan <- as.data.frame(summary(logmodel_fit, pars = c( del3namesstan))$summary)
del3stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu5"))$summary)
delt3stan$parameters <- c(del3namesstan)

@@

# Delta 4 
del4namesstan <- paste("delta4", "[",1:52,"]", sep="")
delt4stan <- as.data.frame(summary(logmodel_fit, pars = c( del4namesstan))$summary)
del4stanmu <- as.data.frame(summary(logmodel_fit, pars = c("mu6"))$summary)
delt4stan$parameters <- c(del4namesstan)

@@

### Obtain all individual estimates delta 
hier_modeld <- logmodel_fit  # model fit

@@

|    In case we have different models, we can compare them to check under which model the data is most likely. This can be done with the Bayes factor. In `rstan` the Bayes factor can be obtained by using the `bridgesampling` package. To be able to use this package, it is important that the `.stan` file is written in a way that it saves the constants (i.e., `target += normal_lpdf(mu4 | a3, sqrt(b3));`).   
|    First, we fit all the models (except for the full model that we already fitted. 
&nbsp;

\scriptsize 
```{r modelcomparison, eval = FALSE, echo = TRUE}
# Fit the other models 
## Null model 
BFnulllogmodel <- stan(file = "myPath/logstanmodelh0.stan", 
                    data = myData_list, 
                    iter = 4000, 
                    chains = 4,  
                    warmup = 1000, 
                    cores = 4)

## Side model 
BFsidelogmodel <- stan(file = "myPath/logstanmodelhside.stan", 
                       data = myData_list, 
                       iter = 4000, 
                       chains = 4, 
                       warmup = 1000, 
                       cores = 4)

## Digit model 
BFdigitlogmodel <- stan(file = "myPath/logstanmodelhdigit.stan", 
                         data = myData_list, 
                         iter = 4000, 
                         chains = 4, 
                         warmup = 1000, 
                         cores = 4)
```
\normalsize

@@

\scriptsize 
```{r bfpart2, eval = FALSE, echo = TRUE}
# Bridge sampling 
# library(bridgesampling)
bridge_Hnulllog <- bridge_sampler(BFnulllogmodel)
bridge_Hsidelog <- bridge_sampler(BFsidelogmodel)
bridge_Hdigitlog <- bridge_sampler(BFdigitlogmodel)
bridge_Hfulllog <- bridge_sampler(logmodel_fit)  # in case you just fitted the model yourself

@@

### Fit model again but do not run any chains 
new_logmod <- stan(file = "myPath/myLogModel.stan", 
                data = myData_list, 
                chains = 0)

### Now do the bridgesampling 
bridge_Hfulllog <- bridge_sampler(logmodel_fit, new_logmod)
```
\normalsize

@@

# Obtain BF (in this way evidence in favor of the null hypothesis)
bridgesampling::bf(bridge_Hnulllog, bridge_Hfulllog)   # Null against full

# Obtain BF (in this way evidence in favor of the full hypothesis)
bridgesampling::bf(bridge_Hfulllog, bridge_Hnulllog)   # Full against null
bridgesampling::bf(bridge_Hfulllog, bridge_Hsidelog)   # Full against side 
bridgesampling::bf(bridge_Hfulllog, bridge_Hdigitlog)  # Full against digit 
```
\normalsize
